---
title: "effects"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file visualises and tabulates parameters from models from the motor slowing reward project and specifically the pupil pilot data.

This is just a demo/example of how you might fit hierarchical gams to pupil data with n=5.

# load the libraries that we will be using #

## install ##

only the additional packages (not already installed are required).
in this case, that is whatever we use to make tables.

```{r install-pkg}
# install.packages(c("flextable"))
```

take a snapshot of loaded packages and update the lock.file using renv

```{r snapshot-renv}
# take a snapshot and update the lock.file
# renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

## load ##

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "brms", 
         "tidybayes", "bayesplot", "future", "parallel", "modelr", "flextable")

lapply(pkg, library, character.only = TRUE)
```

## adjust multicore settings ##

```{r set-options}
options(mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()
```

## plot settings ##

theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18), 
          title = element_text(size = 18),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

## table settings ##

```{r}
set_flextable_defaults(
  theme_fun = theme_vanilla,
  background.color = "white"
)
```


## read in previously saved brms models object files and/or fits ##

If available, this saves time computing them each time 

```{r}
## this was the smooth by pid,condition interaction, which had the best model fit
bp1 <- readRDS("models/bp1.rds")
```

now read in the posterior predictions, if already available.

```{r}

```

## read in the data ##

```{r}
datam <- read_csv("data/datam_n5.csv") %>%
  mutate(pid = factor(pid),
         condition = factor(condition, 
                            levels = c("neutral", "reward")))
head(datam)
str(datam)
```

# section 1 #

## Look at the point and interval estimates for models ##

You can do this, but my understanding is that these don't make much sense on their own.

e.g., the terms in isolation are hard to interpret in a meaningful way. Take a look below.

this is the summary of the model

```{r}
summary(bp1)
```

and these are all of the parameters

```{r}
vars <- as_tibble(get_variables(bp1)) %>%
  rename(term=value) %>% 
  filter(! term %in% c("lprior", "accept_stat__", "treedepth__", "lp__",
                       "divergent__", "n_leapfrog__", "energy__", "stepsize__"),
         str_detect(term, "zs", negate = TRUE)) %>% 
  print(n=Inf)
```

instead you need to take all the parameters and generate posterior predictions because it is the combination of parameters that the model is using to make predictions.

# section 2 #

## posterior predictions by condition ##

let's do this for bp1

```{r}
epred <- datam %>%
  group_by(pid, condition) %>% 
  data_grid(time = seq(15001, 25000, 100)) %>%
  add_epred_draws(bp1) %>% 
head(epred)
```

pred check

```{r}
## check the right number of samples per pid
epred %>%
  ungroup() %>%
  group_by(pid) %>% 
  distinct(time) %>% 
  tally()

## and in total
epred %>%
  ungroup() %>%
  distinct(time) %>% 
  tally()
```

plot

plot by pid

```{r}
p2.1 <- 
  ggplot(data = epred, aes(x = time, y = .epred, 
                                colour=condition, fill = condition)) +
  stat_lineribbon(.width = .95, alpha = 0.7) +
  geom_vline(xintercept = 20000, linetype = "dashed", colour = "darkgrey") +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(breaks=seq(15000,25000,5000)) +
  facet_wrap(~pid, nrow=1) +
  labs(title = "Posterior predictions by participant")
p2.1

ggsave("figures/preds_pid.jpeg",
       width = 10, height = 8)
```

create group data

```{r}
epred_group <- epred %>%
  ungroup() %>% 
  group_by(condition, time, .draw) %>% 
  summarise(group_epred = mean(.epred))
epred_group
```

plot the group data

```{r}
p2.2 <-
  ggplot(data = epred_group, aes(x = time, y = group_epred,
                                colour=condition, fill = condition)) +
  stat_lineribbon(.width = .95, alpha = 0.7) +
  geom_vline(xintercept = 20000, linetype = "dashed", colour = "darkgrey") +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(breaks=seq(15000,25000,5000)) +
  labs(title = "Posterior predictions")
p2.2

ggsave("figures/preds_group.jpeg",
       width = 10, height = 8)
```

## calculate quantile intervals and make a table ##

quantile intervals at the group level (just pick 1s timebins to make it simpler)

```{r}
epred_group_q <- epred_group %>%
  group_by(.draw) %>% 
  slice(seq(1, n(), 10)) %>% 
  mutate(time=factor(time)) %>% 
  group_by(condition, time) %>% 
  median_qi(group_epred)
head(epred_group_q)
```

make a table

```{r}
## make a table using flextable (still needs a title but that can be fixed later, depending on the doc it is being used in)
epred_tab <- epred_group_q %>%
  select(-.width, -.point, -.interval) %>%
  rename(.epred = group_epred) %>% 
  mutate(across(where(is.double), \(x) round(x, 2))) %>% 
  flextable() 
epred_tab

## save it
save_as_image(epred_tab, path = "tables/epred.png")

## this table still needs polishing, but you get the basic idea
```

# section 3 #

## posterior predictions by difference score ##

## calculate difference scores based on condition (reward > neutral) ##

wrangle 

```{r}
epred_diff <- epred %>%
  ungroup() %>% 
  select(pid, condition, time, .draw, .epred) %>% 
  pivot_wider(names_from = "condition",
              values_from = ".epred") %>% 
  mutate(diff = reward - neutral)
head(epred_diff)
```

plot

```{r}
p3.1 <- ggplot(epred_diff, aes(x = time, y = diff)) +  
  stat_lineribbon(.width = .95, alpha = 0.7) +
  geom_vline(xintercept = 20000, linetype = "dashed", colour = "darkgrey") +
  geom_hline(yintercept = 0, linetype = "solid", colour = "darkred") +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(breaks=seq(15000,25000,5000)) +
  theme(legend.position = "none") +
  facet_wrap(~pid, nrow = 1) +
  labs(title = "Posterior predictions (reward > neutral)")
p3.1

ggsave ("figures/preds_diff_pid.jpeg",
        width = 10, height = 8)
```

group diff

```{r}
epred_diff_group <- epred_diff %>% 
  group_by(time, .draw) %>% 
  summarise(group_diff = mean(diff))
epred_diff_group
```

plot

```{r}
p3.2 <- ggplot(data = epred_diff_group, aes(x = time, y = group_diff)) +
  stat_lineribbon(.width = .95, alpha = 0.7) +
  geom_vline(xintercept = 20000, linetype = "dashed", colour = "darkgrey") +
  geom_hline(yintercept = 0, linetype = "solid", colour = "darkred") +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(breaks=seq(15000,25000,5000)) +
  theme(legend.position = "none") +
  labs(title = "Posterior predictions (reward > neutral)")
p3.2

ggsave ("figures/ppreds_diff_group.jpeg",
        width = 10, height = 8)
```


## calculate quantile intervals on the difference score, plot them and make a table ##

(for ease, maybe just pick 20 x axis values - i.e., times, to provide a summary)

at the group level

```{r}
epred_diff_group_q <- epred_diff_group %>%
  group_by(.draw) %>% 
  slice(seq(1, n(), 10)) %>% 
  mutate(time=factor(time)) %>% 
  group_by(time) %>% 
  median_qi(group_diff)
head(epred_diff_group_q)
```

plot using geom_interval (this is less informative than the ribbon plot, but it might be more familiar)

```{r}
p3.3 <- ggplot(data = epred_diff_group_q, aes(x = time, y = group_diff)) +
  geom_pointinterval(aes(ymin=.lower, ymax=.upper),
                     position = position_dodge(width = .7)) +
  geom_hline(yintercept = 0, linetype = "solid", colour = "darkred") +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") 
p3.3

ggsave ("figures/ppreds_diff_interval.jpeg",
        width = 8, height = 4)
```

make a table

```{r}
## make a table using flextable (still needs a title but that can be fixed later, depending on the doc it is being used in)
epred_diff_tab <- epred_diff_group_q %>%
  select(-.width, -.point, -.interval) %>%
  rename(.epred_diff = group_diff) %>% 
  mutate(across(where(is.double), \(x) round(x, 2))) %>% 
  flextable() 
epred_diff_tab

## save it
save_as_image(epred_diff_tab, path = "tables/epred_diff.png")

## this table still needs polishing, but you get the basic idea
```

